Covariant Phase Space via MultiSymplectic Geometry

### 2017-05-15 cpsviams.tex
Commutative Diagram

\usepackage{amsmath}
\usepackage[multisym, geomec,basic]{../Math-Symbols-List/commonmathsymb}
\usepackage{standalone}
\usepackage{graphicx}
\usepackage{wrapfig}

@@

\begin{wrapfigure}{L}{0.41\textwidth}
\centering
\includegraphics[width=0.40\textwidth]{Pictures/jacobi.png}
\caption{\footnotesize \label{fig:jacobi}$J[\phi]$ and $D$ target on the same space.}
\end{wrapfigure}
where:
\begin{itemize}
	\item $\phi_\lambda$ is an arbitrary variation of $\phi_0 = \phi$ generated by $\delta\phi$.
	\item exactly as noted for the tangent vector of $\Conf$, we have for any $x\in M$:
		\begin{displaymath}
			\left.\dfrac{\textrm{d}}{\textrm{d} \lambda} D_\cdot [\phi_\lambda]\right\rvert_{\lambda=0}(x) \in V_M W
			\quad \Rightarrow \quad
			\left.\dfrac{\textrm{d}}{\textrm{d} \lambda} D_\cdot [\phi_\lambda]\right\rvert_{\lambda=0} \in D[\phi]^\ast \left( V_M \left( V^\TwistedLinear \right)\right)
		\end{displaymath}
		because $D_\cdot[\phi_\lambda]$ can be regarded as a section of $V^\TwistedLinear F$ over $M$, 
		thus the latter definition does not depend from the particular choice of variation $\phi_\lambda$ 
		but only from its infinitesimal generator $\delta\phi$.
	\item $\sigma$ is the canonical bundle-morphism
		\begin{displaymath}
			\sigma:  0^\ast \left( V_M W \right)  \; \rightarrow \;  0^\ast \left(V_E W \right) \simeq W
		\end{displaymath}
		 where $V_M W$ and $V_F W$ are the vertical sub-bundle of $T W$ w.r.t. $M$ and $F$ respectively 
		 and $0$ is the zero-section of $W$ over $F$.
\end{itemize}

@@

\section{Construction of $\sigma$}
Consider the two stacked bundle, $ W \rightarrow E$ vector bundle, $E \rightarrow M$ smooth bundle.

@@

We have the following short exact sequence of bundle morphism over $W$
\begin{figure}
  \centering
  \caption{short exact sequence}
\end{figure}
because

@@

The claimed right splitting is the following:
\begin{align*}
		u = T0 \circ m \; : & \; \pi^\ast_{E,W} \left(TE \right) \rightarrow TE \rightarrow TW \\
						& \{w\} \times \{V_{\pi(w)} \} \mapsto V_{\pi(w)} \mapsto \left( T_{\pi(w)}0\right) (V_{\pi(w)})
\end{align*}

@@

\begin{figure}
  \centering
  \includegraphics{Pictures/jacobi.png}
  \caption{.}
\end{figure}

\begin{figure}
  \centering
  \includegraphics{Pictures/jacobi.png}
  \caption{.}
\end{figure}

\begin{figure}
  \centering
  \includegraphics{Pictures/fieldJacobi.png}
  \caption{.}
\end{figure}

\begin{figure}
  \centering
  \includegraphics{Pictures/sigma.png}
  \caption{.}
\end{figure}

>>>>

\usepackage{amsmath}
\usepackage[multisym, geomec,basic]{../Math-Symbols-List/commonmathsymb}
\usepackage[subpreambles=true]{standalone}
\usepackage{graphicx}
\usepackage{wrapfig}

@@

\begin{wrapfigure}{L}{0.41\textwidth}
\centering
\includegraphics[width=0.40\textwidth]{Pictures/Jacobi.png}
\caption{\footnotesize \label{fig:jacobi}$J[\phi]$ and $D$ target on the same space.}
\end{wrapfigure}
where:
\begin{itemize}
	\item $\phi_\lambda$ is an arbitrary variation of $\phi_0 = \phi$ generated by $\delta\phi$.
	\item exactly as noted for the tangent vector of $\Conf$, we have for any $x\in M$:
		\begin{displaymath}
			\left.\dfrac{\textrm{d}}{\textrm{d} \lambda} D_\cdot [\phi_\lambda]\right\rvert_{\lambda=0}(x) \in V_M W
			\quad \Rightarrow \quad
			\left.\dfrac{\textrm{d}}{\textrm{d} \lambda} D_\cdot [\phi_\lambda]\right\rvert_{\lambda=0} \in D[\phi]^\ast \left( V_M \left( V^\TwistedLinear \right)\right)
		\end{displaymath}
		because $D_\cdot[\phi_\lambda]$ can be regarded as a section of $V^\TwistedLinear F$ over $M$, 
		thus the latter definition does not depend from the particular choice of variation $\phi_\lambda$ 
		but only from its infinitesimal generator $\delta\phi$.
	\item $\sigma$ is the canonical bundle-morphism
		\begin{displaymath}
			\sigma:  0^\ast \left( V_M W \right)  \; \rightarrow \;  0^\ast \left(V_E W \right) \simeq W
		\end{displaymath}
		 where $V_M W$ and $V_F W$ are the vertical sub-bundle of $T W$ w.r.t. $M$ and $F$ respectively 
		 and $0$ is the zero-section of $W$ over $F$.
\end{itemize}

@@

\subsection{Construction of $\sigma$}
Consider the two stacked bundle, $ W \rightarrow E$ vector bundle, $E \rightarrow M$ smooth bundle.

@@

We have the following short exact sequence of bundle morphism over $W$
\begin{figure}
  \centering
  \includestandalone[mode=buildnew]{sigma_exact}
\end{figure}
because

@@

The claimed right splitting is the following:
\begin{align*}
		u = T0 \circ m \; : & \; \pi^\ast_{E,W} \left(TE \right) \rightarrow TE \rightarrow TW \\
						& \{w\} \times \{V_{\pi(w)} \} \mapsto V_{\pi(w)} \mapsto \left( T_{\pi(w)}0\right) (V_{\pi(w)})
\end{align*}
In fact, since $0 \in \Gamma (W,E)$:

\begin{displaymath}
	\textrm{T}_e 0 \circ \textrm{T}_\omega \pi_{E,W} V =
	\textrm{T}_\omega \left( 0 \circ \pi_{E,W} \right) V = 
	V \qquad \forall V \in \left( V_M W \right)_\omega ,\; \omega \in \pi^-1(e)
\end{displaymath}
Therefore, we have the following decomposition:

\begin{displaymath}
	\left( V_M W \right)_{0(e)} = 
	i\left[ V_E W \right]_{0(e)} \oplus \left( T_e 0 \left[ V_e E\right] \right)
\end{displaymath}
determining an intrinsic mapping 
\begin{displaymath}
	\sigma_e : \left( V_M W \right)_{0(e)} \rightarrow \left( V_E W \right)_{0(e)} \simeq W_e
\end{displaymath}

which is simply a  projection on the first component.

Since we have such projection for all $e \in E$ and it acts fiberwise, it can be regarded as a bundle morphism over $E$:
\begin{displaymath}
	\sigma : 0^\ast \left( V_M W \right) \rightarrow 0^\ast \left( V_E W \right) \simeq W
\end{displaymath}

\begin{wrapfigure}{L}{0.41\textwidth}
\centering
\includegraphics[width=0.40\textwidth]{Pictures/sigma.png}
  %\caption{.}
\end{wrapfigure}
Visually this is a map which "verticalize" w.r.t. $E$ the vertical vectors w.r.t. $M$ along $0(E) \subset W$

\begin{figure}
  \centering
  \includegraphics[width=0.41\textwidth]{Pictures/fieldJacobi.png}
  \caption{.}
\end{figure}

\newpage
\section{Covariant approach: pre-symplectic form}
\begin{wrapfigure}{L}{0.41\textwidth}
\centering
\includegraphics[width=0.40\textwidth]{Pictures/symform.png}
  \caption{(temp)}
\end{wrapfigure}

Rough strategy:
\begin{itemize}
	\item Find a way to contract $\delta \phi$ and $\delta \phi '$ , seen as tangent vector field on $E$, with the multi-symplectic forms $\omega$.
	\item pull-back the obtained $m-1$ form to $M$, oriented manifold.
	\item Integrate thus form on a suitable $(m-1)$ dimensional sub-manifold of $M$.
	\item take the resulting number as the evaluation of the desired form on the two variations.  
\end{itemize}

Rigourosly, we define $\forall \phi \in \Conf$, $\forall \Sigma \subset M$ $m-1$ dimensional, compact submanifold

\begin{definition}[(covariant) pre-symplectic form in $\phi$ pertaining to $\Sigma$]
	\begin{displaymath}
		\Omega_\Gamma ( \phi ) : \textrm{T}_\phi \Conf \times \textrm{T}_\phi \Conf \rightarrow \Real
\end{displaymath}		
	S.t:
	\begin{align*}
		\left\langle	\Omega_\Sigma ( \varphi) \right.\left\lvert \delta \varphi_1 , \delta \varphi_2 \right\rangle &=		
		\int_\Gamma (j^1 \varphi)^\ast \left[ i_{\delta\varphi_2} i_{\delta \varphi_1} \omega_\Lagrangian \right]
		& \qquad \textrm{(Lagrangian framework)} \\
		\left\langle	\Omega_\Sigma ( \varphi, \pi) \right.\left\lvert ( \delta \varphi_1, \delta \pi_1),
		( \delta \varphi_2, \delta \pi_2) \right\rangle &=		
		\int_\Gamma (\varphi, \pi)^\ast \left[ i_{( \delta \varphi_2, \delta \pi_2)} i_{( \delta \varphi_1, \delta \pi_1)} \omega_\Hamiltonian \right]
		& \qquad \textrm{(Hamiltonian framework)} 		
	\end{align*}
\end{definition}

We can easily recognize that at this level $J^1 E$ and $\LinDualJet$ takes a similar role

@@
